\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage[backend=biber,style=numeric-comp,sorting=none]{biblatex}

\addbibresource{references.bib}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=blue!50!black,
  urlcolor=blue!50!black,
  pdftitle={Verifiable Cross-Chain Risk Intelligence: OmniRisk, RayaChain, and Bittensor},
  pdfauthor={OmniRisk Research},
}

\title{Verifiable Cross-Chain Risk Intelligence:\\
       An Agentic Architecture over LayerZero V2 and Bittensor}
\author{OmniRisk Research \\ \texttt{research@omnirisk.io}}
\date{May 2026 \\ \textbf{Working paper, v0.5}}

\begin{document}
\maketitle

\begin{abstract}
Cross-chain bridges have become the single largest concentrated source of value loss in
decentralised finance: industry trackers attribute roughly USD~2B of theft in the first
eight months of 2022 alone to bridge attacks~\cite{chainalysis2022bridges}, and the most
recent peer-reviewed systematisation places cumulative losses at approximately
USD~3.1B~\cite{augusto2024sok}. The risk telemetry that participants need to navigate
this environment remains \emph{siloed per chain} and \emph{non-verifiable}: there is no
widely deployed mechanism that lets a downstream user confirm that a route- or
token-level risk signal was produced honestly. This paper asks a single research
question: \emph{can cross-chain risk intelligence be made verifiable through
Bittensor-style decentralised validation, and at what cost in latency, calibration, and
adversarial robustness?} Our central claim is that the answer is yes, and that the
correct primitive is not yet another risk API but a \textbf{verification architecture in
which route-level risk outputs are economically evaluated by decentralised validators}.
We present \textsc{OmniRisk}, a deployed agentic risk-intelligence layer that ingests
cross-chain flow data from \textsc{RayaChain}, an Omnichain Fungible Token (OFT)
deployment on LayerZero~V2~\cite{zarick2024layerzero,layerzero2024oft}, and emits
structured route and token risk signals. The architecture splits responsibility into
(i)~OmniRisk as the intelligence layer, (ii)~RayaChain as the cross-chain data and
execution layer, and (iii)~Bittensor's Yuma Consensus~\cite{rao2021bittensor,
steeves2022incentivizing,opentensor2024yuma} as the decentralised verification layer
for the intelligence itself. We argue that anchoring the verification layer on the
Subtensor EVM (chain~ID~945)~\cite{opentensor2024evm} is the natural fit because it is
the only production EVM whose validator set is already economically aligned to scoring
the quality of AI outputs. The paper formalises the route, risk-score, and validator-loss
primitives, specifies the miner and validator protocols required to make this work,
situates the design against the cross-chain bridge security and decentralised-ML
verification literatures~\cite{augusto2024sok,zhang2023sok,wu2025safeguarding,
conway2024opml,ganescu2024trust}, and lays out the evaluation framework --- including
the four baselines required to support a publication-grade empirical claim --- that
will be applied to historical RayaChain flows.
\end{abstract}

\section{Introduction}\label{sec:intro}

\textsc{OmniRisk} is a live cross-chain risk intelligence service, and \textsc{RayaChain}
is an actively deploying cross-chain transfer fabric. RayaChain's transfer lanes are
implemented as Omnichain Fungible Tokens (OFTs) on LayerZero~V2, the protocol family
that separates message verification (Decentralised Verifier Networks, DVNs) from message
execution (Executors) and exposes per-application security configuration as a
permissionless surface~\cite{zarick2024layerzero,layerzero2024oft}. As of submission,
RayaChain is operating across seven EVM and non-EVM ecosystems --- Ethereum, Arbitrum,
Base, BNB Smart Chain, Polygon, Optimism, and Solana --- and is scoping an eighth lane
into the Subtensor EVM (chain~ID~945)~\cite{opentensor2024evm}.

The motivation for the eighth lane is not capacity but \emph{verifiability}. Routing
value across heterogeneous chains exposes participants to a class of risks that survey
work has documented in detail: design flaws in bridge architectures map onto a stable
catalogue of attack surfaces~\cite{zhang2023sok}, attack-transaction analyses confirm
that the same patterns recur across dozens of incidents~\cite{wu2025safeguarding}, and
losses concentrate in the bridge layer to a degree that no other DeFi primitive
matches~\cite{chainalysis2022bridges,augusto2024sok}. Existing risk tools react to this
by emitting black-box risk scores: a downstream contract or wallet must trust the
operator's API, with no cryptographic or economic mechanism that constrains the
operator to be honest. This is the same primitive failure that gave rise to the
oracle-design literature in the early DeFi era, and we argue it now calls for a
verification layer that is itself decentralised.

\paragraph{Research question.} The paper is organised around one question:
\begin{quote}
\emph{Can cross-chain risk intelligence be made verifiable through Bittensor-style
decentralised validation?}
\end{quote}
This is a hypothesis, not a result. The architecture, the formal definitions
(Section~\ref{sec:formal}), and the evaluation framework (Section~\ref{sec:eval}) are
the apparatus needed to answer it; the answer itself is deferred to follow-up empirical
work. We make the hypothesis falsifiable by stating two specific research questions and
two hypotheses in Section~\ref{sec:rq}.

\paragraph{Novelty claim.} The strongest novelty is not ``agentic AI'' or ``cross-chain
data'' considered separately. The contribution is the \textbf{verification architecture}:
route-level risk outputs are produced by a heterogeneous population of miners and
\emph{economically evaluated} by decentralised validators, with rewards proportional to
agreement with realised cross-chain events. This is, to our knowledge, the first such
construction for cross-chain risk; we situate it precisely against prior work in
Section~\ref{sec:related}.

\paragraph{Contributions.} Concretely, the paper makes five contributions.
\begin{enumerate}
  \item \textbf{An architectural separation of concerns}: OmniRisk = intelligence,
        RayaChain = data and execution, Bittensor = verification (Section~\ref{sec:arch}).
  \item \textbf{A formalisation of the cross-chain risk-scoring task} in terms of
        routes $q$, the risk-score function $s(q,t)$, and a validator loss $L_i$ that
        decomposes into accuracy, consistency, and calibration terms
        (Section~\ref{sec:formal}).
  \item \textbf{A miner protocol specification} for thin OmniRisk-API wrappers that meet
        the determinism and structured-reasoning requirements of Yuma
        scoring~\cite{opentensor2024yuma} (Section~\ref{sec:miner}).
  \item \textbf{A validator-side scoring rule} that evaluates miner outputs against
        ground-truth realised cross-chain risk events, expressed in the Yuma weight-vector
        format with precise definitions of \emph{realised loss}, \emph{material chain-state
        change}, and \emph{confidence calibration} (Section~\ref{sec:validator}).
  \item \textbf{An evaluation framework with explicit baselines} grounded in replayed
        historical RayaChain flows, with an honest discussion of what we cannot yet
        prove (Section~\ref{sec:eval} and Section~\ref{sec:limitations}).
\end{enumerate}

This is a working paper. We are explicit throughout about which components are deployed
today, which are in active integration, and which are design proposals (see
Table~\ref{tab:status} in Section~\ref{sec:status}).

\section{Problem Definition}\label{sec:problem}

\paragraph{Unified risk visibility across chains is missing.} A token's risk profile is
not a property of any single chain. The same OFT can have safe liquidity on Ethereum
and dangerously thin liquidity on a destination rollup; a user routing through it will
realise the worse of the two. Empirical AMM-level evidence shows that liquidity is
materially mis-allocated across L1 and L2s relative to LP-return
equilibrium~\cite{gogol2024liquidity}, which means \emph{route choice} --- not just
token choice --- is a first-order risk decision.

\paragraph{Real-time route risk awareness is missing.} The bridge security literature
documents 49~attack incidents on cross-chain bridges between June~2021 and
September~2024~\cite{wu2025safeguarding}, and a security/privacy taxonomy spanning
57~interoperability solutions identifies recurring failure modes across them
all~\cite{augusto2024sok}. Yet at the moment a user submits a transfer, they typically
have no signal that the lane they are about to use shares its design with a recently
exploited one.

\paragraph{Verifiable intelligence is missing.} Existing route-risk APIs are
\emph{black-box}: a downstream contract must trust that the API operator scored honestly.
Recent work on optimistic and zero-knowledge ML on
blockchain~\cite{conway2024opml,ganescu2024trust} and on multichain ML-based fraud
detection~\cite{palaiokrassas2023multichain} addresses parts of this gap, but no system
that we are aware of combines (a)~live cross-chain risk scoring with (b)~a decentralised
verification layer for the scores themselves.

\section{Research Questions and Hypotheses}\label{sec:rq}

We make the architectural claim of Section~\ref{sec:intro} falsifiable by reducing it to
two research questions and two hypotheses. Each is operationalised by a quantity that the
evaluation framework of Section~\ref{sec:eval} produces.

\paragraph{RQ1.} Does cross-chain route-level telemetry --- the joint state of source-chain
liquidity, destination-chain liquidity, and the OFT lane configuration --- improve risk
prediction relative to single-chain risk features for the same route?

\paragraph{RQ2.} Can decentralised validators reliably distinguish calibrated risk
outputs from uncalibrated or adversarial miner outputs over repeated replay evaluation?
That is: under the validator scoring rule of Section~\ref{sec:validator}, does the Yuma
emission share converge onto miners whose realised-event accuracy and confidence
calibration are independently good?

\paragraph{H1.} Cross-chain liquidity-fragmentation features improve the early-detection
lead time (number of blocks between threshold crossing and the realised event) for
adverse route events, relative to a single-chain liquidity-only model. This hypothesis is
falsified if the lead-time distribution under the cross-chain feature set is statistically
indistinguishable from the lead-time distribution under the single-chain baseline.

\paragraph{H2.} Confidence-calibrated miners receive monotonically higher validator
weights than uncalibrated miners over repeated replay evaluation. That is: holding the
realised-event accuracy term constant, miners whose Brier score on $(s, c)$ is lower
must achieve a strictly higher Yuma reward share. This hypothesis is falsified if the
correlation is non-significant or has the wrong sign.

The deferred empirical paper will report on RQ1 and RQ2 directly and will accept or reject
H1 and H2 against the baselines specified in Section~\ref{sec:eval:baselines}.

\section{Related Work}\label{sec:related}

We organise the literature into six sub-areas, distinguish what each contributes, and
identify the gap that this paper addresses.

\subsection{Cross-chain bridge security surveys}\label{sec:related:bridge-sok}

Two systematisations frame the bridge-security landscape. \cite{augusto2024sok} is the
peer-reviewed SoK at IEEE S\&P 2024: it surveys 212 documents (58 peer-reviewed and 154
grey), enumerates a security/privacy taxonomy across 57 interoperability solutions, and
reports cumulative cross-chain hack losses of approximately USD~3.1B. \cite{zhang2023sok}
provides a complementary attack-surface taxonomy organised by bridge architecture
(lock-and-mint, burn-and-mint, atomic swap, light-client) and identifies recurring
defence gaps. Both are descriptive; neither proposes a verifiable risk-scoring primitive.

\subsection{Bridge attack-detection systems}\label{sec:related:bridge-detection}

\cite{wu2025safeguarding} introduces \textsc{BridgeGuard}, a two-stage graph-mining
detector that combines a global graph view of cross-chain transactions with a local
motif-extraction step. The labelled corpus comprises 203 attack transactions across
49 cross-chain bridge incidents (June~2021 -- September~2024) against 40{,}000 normal
cross-chain transactions, an empirical class-imbalance ratio of approximately
$1{:}197$; the authors report a recall improvement of 36.32 percentage points over
prior state-of-the-art tools and the ability to flag previously unseen attack
transactions~\cite{wu2025safeguarding}. \textsc{BridgeGuard} is the closest existing
analogue to one component of OmniRisk's realised-event channel and we use its labelled
corpus as the canonical $E4$ exploit-cluster label source in
Section~\ref{sec:formal:realised}. Two differences from our work are essential:
(i)~it is a single-operator detector with no verification layer over its outputs, and
(ii)~it is retrospective event-detection rather than ex-ante route-risk prediction.
A consequence for our evaluation framework (Section~\ref{sec:eval}) is that the
$1{:}197$ imbalance is the natural prior for the realised-event channel, which forces
the baselines of Section~\ref{sec:eval:baselines} to be evaluated on imbalance-aware
metrics (precision-recall AUC, not just ROC-AUC) and motivates threshold selection
that prioritises recall on the rare positive class.

\subsection{Multichain fraud detection}\label{sec:related:multichain}

\cite{palaiokrassas2023multichain} releases a labelled multichain fraud-detection dataset
spanning 23 DeFi protocols and 12 chains, with baseline gradient-boosted and deep
classifiers. The contribution is the dataset and baselines, not a route-level risk model
or a verification mechanism. We adopt their multichain framing and note that route-level
telemetry (the OFT lane and its DVN configuration) is an additional feature class not
present in their corpus.

\subsection{Decentralised ML verification}\label{sec:related:dml}

Two complementary verification primitives have appeared in 2024.
\cite{conway2024opml} (\textsc{opML}) verifies off-chain ML inference via an
optimistic-rollup-style fraud-proof challenge game on chain. \cite{ganescu2024trust}
(\textsc{snarkGPT}) verifies transformer inference via zero-knowledge succinct
arguments. Both are general-purpose ML-verification primitives; neither targets
cross-chain risk scoring specifically and neither relies on an economic
validator-quorum mechanism. Bittensor's Yuma consensus is a third verification
primitive --- an economic-validator-quorum mechanism --- and our contribution is to apply
it to the cross-chain risk-scoring task.

\subsection{Oracle and data-verification mechanisms}\label{sec:related:oracle}

The broader DeFi systematisation~\cite{werner2022sok} catalogues data-feed and oracle
mechanisms as a core DeFi primitive: trusted-feeder oracles, Schelling-point oracles, and
zk-attested feeds. The route-risk primitive proposed in this paper is structurally a
\emph{decentralised attestation feed}: it makes a claim of the form ``route $q$ has risk
$s$ at time $t$'', and that claim is meant to be consumed by downstream contracts. The
gap is that existing oracle taxonomies cover \emph{prices} (sourced from venues) and
\emph{external events} (sourced from APIs), but not \emph{model outputs} produced by an
agentic risk pipeline. The Bittensor verification layer of Section~\ref{sec:arch:bittensor}
is the first economic-validator-quorum primitive deployed at production scale that
naturally accommodates model-output attestations of this form.

\subsection{Bittensor incentive mechanisms}\label{sec:related:bittensor}

The Bittensor incentive mechanism is specified canonically in
\cite{rao2021bittensor} and analysed formally in \cite{steeves2022incentivizing}, which
proves collusion-resistance up to coalitions controlling less than 50\% of stake-weighted
rank. The current production specification is documented in
\cite{opentensor2024yuma}, with the Subtensor EVM execution environment in
\cite{opentensor2024evm}. \cite{lui2025bittensor} is the most recent critical empirical
analysis: it documents real centralisation pressure under the dynamic-TAO regime and
should be read alongside the formal collusion-resistance bound rather than as a
contradiction of it. We treat the formal bound as a property of the algorithm and the
empirical findings as a property of the deployed network's stake distribution; both
matter for the security analysis of Section~\ref{sec:security}.

\subsection{Where the gap is}\label{sec:related:gap}

Existing work covers (a)~bridge security descriptively, (b)~retrospective attack
detection, (c)~multichain fraud-detection datasets, (d)~general-purpose ML-verification
primitives, (e)~oracle taxonomies for prices and events, and (f)~the Bittensor incentive
mechanism in isolation. No published system simultaneously
(i)~scores route-level cross-chain risk in real time,
(ii)~submits those scores to a decentralised validator quorum that scores them against
realised events, and
(iii)~anchors the resulting consensus on chain so it can be consumed verifiably by
downstream contracts. That intersection is what this paper proposes.

\section{System Architecture}\label{sec:arch}

We split responsibility across three layers; the dataflow is summarised in
Figure~\ref{fig:arch-text}.

\begin{figure}[h]
\centering
\begin{small}
\begin{tabular}{l}
\toprule
\textbf{Layer 1 --- Intelligence (\textsc{OmniRisk})} \\
~~observes: chain heads, OFT \texttt{UlnConfig}/\texttt{ExecutorConfig}, AMM depth,
incident-cluster proximity \\
~~emits: $(s_i, c_i, r_i)$ per route query $q$ (Section~\ref{sec:miner}) \\
\midrule
\textbf{Layer 2 --- Data \& Execution (\textsc{RayaChain} on LayerZero V2)} \\
~~7 live lanes: Ethereum, Arbitrum, Base, BSC, Polygon, Optimism, Solana \\
~~+1 planned lane: Subtensor EVM (chain ID 945) \\
~~publishes: \texttt{OFTSent}/\texttt{OFTReceived} events, lane configuration, latency \\
\midrule
\textbf{Layer 3 --- Verification (\textsc{Bittensor} subnet, Yuma Consensus)} \\
~~validators: score miner outputs against realised events (Section~\ref{sec:validator}) \\
~~Yuma rule: clip + aggregate via Eq.~\eqref{eq:yuma-consensus};
emit miners share $M_j$ \\
~~anchored on: Subtensor EVM as the score-commitment surface
(Section~\ref{sec:subtensor}) \\
\bottomrule
\end{tabular}
\end{small}
\caption{Layered dataflow. Each layer is independently failure-isolated: an outage in
Layer~1 stops new scores but does not corrupt anchored history; an outage in Layer~3
stops verification but does not stop transfers in Layer~2.}
\label{fig:arch-text}
\end{figure}

\subsection{OmniRisk --- Agentic Intelligence Layer}\label{sec:arch:omnirisk}

OmniRisk is a deployed system that observes cross-chain liquidity health, scores route
risk, and surfaces anomalous on-chain activity. We describe it as \emph{agentic} in the
narrow sense used in recent LLM systems work: a control loop that ingests chain state,
selects which downstream tools (DEX quote APIs, on-chain analytics endpoints,
historical incident databases) to consult, and emits a structured response with an
explicit confidence and a chain-of-evidence. The agentic framing is contrasted with
\emph{static} analytics --- precomputed dashboards, fixed risk-scoring formulas --- which
do not adapt to the question being asked.

\subsection{RayaChain --- Execution and Data Layer}\label{sec:arch:rayachain}

RayaChain is the OFT lane fabric. Each lane is implemented as either a standard OFT
(burn on source, mint on destination) or an OFT Adapter (lock on source, mint on
destination) where a legacy ERC-20 already exists, exactly as specified in the OFT
standard~\cite{layerzero2024oft}. RayaChain inherits LayerZero~V2's permissionless
security configuration: each lane chooses its own DVN set and execution path, which is
the surface OmniRisk monitors when scoring lane-level risk~\cite{zarick2024layerzero}.

The current deployment covers seven chains (Ethereum, Arbitrum, Base, BNB Smart Chain,
Polygon, Optimism, Solana). The proposed eighth lane is the Subtensor EVM
(chain~ID~945)~\cite{opentensor2024evm}, motivated in
Section~\ref{sec:arch:bittensor}.

\subsection{Bittensor --- Verification Layer}\label{sec:arch:bittensor}

Bittensor is a decentralised network in which \emph{miners} produce outputs for a
specific task and \emph{validators} score them; rewards are emitted via Yuma Consensus
on the Subtensor blockchain~\cite{rao2021bittensor,steeves2022incentivizing,
opentensor2024yuma}. We summarise the mechanics here because the validator scoring rule
of Section~\ref{sec:validator} composes with them.

Let $V$ index subnet validators, $M$ index subnet miners, and $S_i$ denote the stake of
validator $i \in V$. Each validator submits a weight $W_{ij} \in [0,1]$ for each miner
$j$. Yuma computes a benchmark consensus weight per miner as
\begin{equation}
  \overline{W}_j \;=\; \arg\max_{w \in [0,1]} \;
    \Bigl\{\, \textstyle \sum_{i \in V} S_i \cdot \mathbf{1}\!\left[W_{ij} \ge w\right]
    \;\ge\; \kappa \cdot \textstyle \sum_{i \in V} S_i \,\Bigr\}
  \label{eq:yuma-consensus}
\end{equation}
with $\kappa = 0.5$ (the default consensus parameter): $\overline{W}_j$ is the largest
weight that at least the $\kappa$-stake-fraction of validators is willing to assign to
miner $j$. Each validator's submitted weight is then clipped against this benchmark,
$\widehat{W}_{ij} = \min(W_{ij}, \overline{W}_j)$, which bounds the influence of any
single validator above the consensus. Aggregate miner ranking is
$R_j = \sum_{i \in V} S_i \cdot \overline{W}_{ij}$ and the per-epoch miner emission share
is $M_j = R_j / \sum_{k \in M} R_k$~\cite{opentensor2024yuma}. Validators in turn earn
through a bond/dividend matrix $B_{ij}$ that is updated as an EMA of the
stake-weighted clipped weights, so validators are rewarded for being early to the
eventual consensus rather than for following it~\cite{opentensor2024yuma,
steeves2022incentivizing}. The mechanism is provably collusion-resistant up to
coalitions controlling less than 50\% of stake-weighted
rank~\cite{steeves2022incentivizing}.

The implication for a cross-chain risk subnet is concrete: a miner is only economically
viable if its risk scores fall within the consensus envelope of the validator quorum;
miners that deviate are clipped out of the emission, and validators that consistently
support deviating miners lose bonded yield. This is the property we want from a risk
oracle, and it is not provided by any centralised risk API. The Subtensor EVM
(chain~ID~945) gives us an Ethereum-compatible execution environment in which to anchor
commitments to score outputs without leaving the Bittensor trust
domain~\cite{opentensor2024evm}.

\section{OmniRisk--Bittensor Integration Model}\label{sec:integration}

The functional mapping is:
\begin{itemize}
  \item \textbf{OmniRisk} generates risk scores and anomaly signals (intelligence).
  \item \textbf{Bittensor validators} act as independent risk-verification nodes
        (verification).
  \item \textbf{RayaChain} contributes the cross-chain flow data and is also one of
        the execution venues whose lanes are being scored (data + execution).
\end{itemize}

A subnet specialised for cross-chain financial intelligence exposes a single task to
its miners: given a route specification (source chain, destination chain, OFT, size,
deadline), return a structured risk score. We do not require miners to all use the same
internal model --- the discipline comes from the validator scoring rule
(Section~\ref{sec:validator}), not from the miner implementation.

\section{Formal Definitions}\label{sec:formal}

We fix notation that the miner specification (Section~\ref{sec:miner}) and the validator
specification (Section~\ref{sec:validator}) both refer to.

\subsection{Route}\label{sec:formal:route}

A \emph{route} is a tuple
\begin{equation}
  q \;=\; (c_s,\; c_d,\; a,\; x,\; t)
  \label{eq:route}
\end{equation}
where $c_s$ is the source chain identifier, $c_d$ the destination chain identifier, $a$
the asset (an OFT in the RayaChain deployment), $x \in \mathbb{R}_{>0}$ the transfer
size denominated in $a$, and $t$ the query time (a wall-clock instant; in
implementation, the most recent finalised block on each of $c_s$ and $c_d$, paired). The
chain identifiers are drawn from the set of RayaChain-active chains
\(
  \mathcal{C} = \{\text{Ethereum},\; \text{Arbitrum},\; \text{Base},\; \text{BNB},\;
  \text{Polygon},\; \text{Optimism},\; \text{Solana}\}
\)
plus the proposed $c = 945$ Subtensor EVM lane (Section~\ref{sec:subtensor}).

\subsection{Risk score}\label{sec:formal:score}

Let $E_{q, t+\Delta} \in \{0, 1\}$ be the indicator of a \emph{realised adverse event}
on route $q$ over the horizon $[t, t + \Delta]$ (the precise event definition is
operationalised in Section~\ref{sec:formal:realised} below). Let $X_{q, t}$ denote the
information set observable on chain at time $t$ that is relevant to $q$ (source/destination
liquidity snapshots, recent OFT throughput, the lane's DVN configuration, recent
incident-cluster proximity).

The route risk score is
\begin{equation}
  s(q, t) \;=\; \Pr\bigl(\, E_{q, t+\Delta} = 1 \;\bigm|\; X_{q, t} \,\bigr) \,.
  \label{eq:risk}
\end{equation}
A miner $i$ emits an estimate $s_i(q, t) \in [0,1]$ of this quantity, together with a
self-reported confidence $c_i(q, t) \in [0,1]$ and a structured reasoning record
$r_i(q, t) \in \mathcal{R}$ (Section~\ref{sec:miner}). Throughout the paper, $\Delta$ is
a subnet-level constant; the deployed value is published with the subnet specification.

\subsection{Realised loss}\label{sec:formal:realised}

The $E_{q, t+\Delta}$ indicator is the disjunction of four observable channels, each
with a precise definition; we explicitly avoid the vague phrase ``something went wrong''.
\begin{description}
  \item[E1: Failed execution.] The OFT message originating from $q$ on $c_s$ either
        reverts on $c_d$ or expires unredeemed past the lane's nonce timeout.
        Operationalised as: zero $\textsc{OFTReceived}$ event on $c_d$ within the
        protocol-defined message-window for the corresponding $\textsc{OFTSent}$ event on
        $c_s$.
  \item[E2: Abnormal slippage.] The realised execution price on $c_d$ deviates from the
        median quoted price at $t$ by more than a threshold $\tau_{\text{slip}}$ (e.g.\
        50 basis points for stablecoin OFTs, 200 bps for volatile OFTs).
  \item[E3: Bridge / lane delay.] The end-to-end latency between $\textsc{OFTSent}$ on
        $c_s$ and $\textsc{OFTReceived}$ on $c_d$ exceeds a per-lane latency threshold
        $\tau_{\text{delay}}$, calibrated as the historical 99th percentile of that lane.
  \item[E4: Exploit-cluster association.] The destination address, the lane, or the OFT
        is identified within $\Delta'$ blocks (typically $\Delta' \le \Delta$) as
        belonging to an attack cluster by an external label source --- in our setup, the
        \textsc{BridgeGuard} attack-transaction corpus~\cite{wu2025safeguarding}
        (49~incidents, 203~attack transactions) and on-chain forensic feeds.
\end{description}
We define $E_{q, t+\Delta} = 1$ if any of $E1$--$E4$ fires; otherwise $E_{q, t+\Delta} =
0$. The precise channel that fired is preserved as auxiliary metadata so a follow-up
analysis can decompose miner accuracy by event type. \emph{Realised loss} in this paper
always refers to this conjunction of observable channels --- never to ``post-route
drawdown'' or other downstream consequences that depend on a user's holding decision.

\subsection{Material chain-state change}\label{sec:formal:material}

The validator consistency rule (Section~\ref{sec:validator}) compares miner outputs at
nearby times $t$ and $t + \delta$. We define a \emph{material chain-state change} on
route $q$ between $t$ and $t + \delta$ as the disjunction:
\begin{enumerate}
  \item the source-chain or destination-chain head changed by at least one block;
  \item the lane's send-side \texttt{UlnConfig} or \texttt{ExecutorConfig} changed
        (specifically: \texttt{confirmations}, \texttt{requiredDVNCount},
        \texttt{requiredDVNs[]}, \texttt{optionalDVNCount}, \texttt{optionalDVNs[]},
        \texttt{optionalDVNThreshold}, \texttt{maxMessageSize}, or
        \texttt{executor})~\cite{layerzero2024oft,zarick2024layerzero}, or the
        receive-side \texttt{UlnConfig} that the destination endpoint enforces;
  \item liquidity depth on either side of the route changed by more than a fraction
        $\eta$ of pre-trade depth (default $\eta = 0.05$);
  \item the OFT contract on $c_s$ or $c_d$ was upgraded or paused.
\end{enumerate}
If none of these holds between $t$ and $t + \delta$, the chain state is \emph{materially
unchanged} and miner responses are required to be byte-identical (modulo the
quantisation precision of Section~\ref{sec:miner}). If any holds, divergent miner
responses are not penalised on the consistency channel.

The lane-configuration coordinates listed in (2) are the LayerZero V2 surface that
OmniRisk monitors per route. The receive-side \texttt{UlnConfig} is what the destination
endpoint actually enforces, while the send-side configuration determines the verification
path the message will follow~\cite{zarick2024layerzero}; both must be tracked because a
mismatch between them blocks delivery and is itself a risk signal.

\subsection{Confidence calibration}\label{sec:formal:calibration}

A miner's confidence $c_i(q, t)$ is \emph{calibrated} if, over a window of $N$ scored
queries with realised outcomes $y_n \in \{0,1\}$, the empirical Brier score
\begin{equation}
  \text{Brier}_i \;=\; \frac{1}{N} \sum_{n=1}^{N} \bigl( s_i(q_n, t_n) - y_n \bigr)^{2}
  \label{eq:brier}
\end{equation}
is no worse than that of a calibrated reference predictor on the same window, and the
self-reported confidence $c_i$ tracks miner accuracy in the sense that the reliability
diagram (binned $c_i$ vs.\ realised accuracy) lies within a tolerance band of the
identity. A miner that reports $c_i = 1$ on every query is automatically miscalibrated
unless its $\text{Brier}_i$ is uniformly zero, which empirically does not occur.

\subsection{Validator loss}\label{sec:formal:vloss}

Let $\mathbf{s}_i, \mathbf{c}_i, \mathbf{y}$ be the miner-$i$ score vector, confidence
vector, and realised-outcome vector over a window of $N$ scored queries. The validator
loss for miner $i$ is the weighted decomposition
\begin{equation}
  L_i \;=\; \alpha \cdot \text{Brier}(\mathbf{s}_i, \mathbf{y})
        \;+\; \beta \cdot \text{Inconsistency}_i
        \;+\; \gamma \cdot \text{Miscalibration}_i
  \label{eq:vloss}
\end{equation}
with non-negative weights $\alpha + \beta + \gamma = 1$ that the subnet publishes as
hyperparameters. The three terms are:
\begin{itemize}
  \item $\text{Brier}(\mathbf{s}_i, \mathbf{y})$ as in Eq.~\eqref{eq:brier} (realised-event
        accuracy);
  \item $\text{Inconsistency}_i$, the fraction of paired queries $(q, t), (q, t+\delta)$
        with no material chain-state change (Section~\ref{sec:formal:material}) for which
        miner $i$ returned different responses;
  \item $\text{Miscalibration}_i$, the expected calibration error
        $\sum_b w_b \,|\, \overline{c}_b - \overline{a}_b \,|$ over equal-mass bins of
        $c_i$, where $\overline{c}_b$ is the mean confidence in bin $b$ and
        $\overline{a}_b$ is the realised accuracy.
\end{itemize}
A validator transforms its loss vector $\{L_i\}_{i \in M}$ into a Yuma weight vector
$\{W_{ij}\}_{j \in M}$ (using the indexing of Section~\ref{sec:arch:bittensor}, where $i$
now indexes the validator and $j$ indexes the miner). We use the
temperature-scaled-softmax form
\begin{equation}
  W_{ij} \;=\; \frac{\exp(-\lambda \, L_j)}{\sum_{k \in M} \exp(-\lambda \, L_k)}
  \label{eq:vweight}
\end{equation}
with $\lambda > 0$ a subnet-published hyperparameter; any monotone decreasing transform
of $L_j$ suffices, and we use the softmax form for numerical stability. The submitted
weights $W_{ij}$ are then aggregated on chain via Eq.~\eqref{eq:yuma-consensus}: a
miner's emission share $M_j$ is determined by the consensus-clipped
$\overline{W}_{ij}$, not by the validator's pre-clip submission. In particular, a
validator that scores miners outside the $\kappa$-stake-fraction consensus envelope
contributes nothing additional to those miners' emissions and erodes its own bonded
yield --- which is what makes the validator scoring rule of this section economically
self-correcting.

\subsection{Worked example}\label{sec:formal:example}

To make the abstraction concrete, consider a single representative route:
\[
  q^{\star} \;=\; (\,c_s = \text{Arbitrum},\; c_d = \text{Base},\; a = \text{USDC OFT},\;
                   x = 100{,}000\;\text{USDC},\; t \text{ = current finalised heads}\,) \,.
\]
The information set $X_{q^{\star}, t}$ that an OmniRisk-backed miner consults contains:
the source-chain USDC liquidity depth observable on Arbitrum DEXes; the destination-chain
liquidity depth on Base; the lane's send-side \texttt{UlnConfig}
(\texttt{requiredDVNCount}, \texttt{requiredDVNs[]}, \texttt{optionalDVNCount},
\texttt{optionalDVNs[]}, \texttt{optionalDVNThreshold}, \texttt{confirmations}); the
\texttt{ExecutorConfig} (\texttt{maxMessageSize}, \texttt{executor}); the receive-side
\texttt{UlnConfig} on Base; rolling latency and message-failure counts on this lane; and
the BridgeGuard incident-cluster proximity for the destination address and the OFT
contract.

A miner returns $(s_i, c_i, r_i)$. Suppose $s_i = 0.07$ (low risk) and $c_i = 0.85$.
The validator subsequently observes the realised outcome over $[t, t + \Delta]$:
suppose the message commits within the lane's 99th-percentile latency
(no $E3$), executes successfully on Base (no $E1$), at within-50-bp slippage
(no $E2$), and the destination address does not appear in any BridgeGuard cluster
within $\Delta'$ blocks (no $E4$). Then $E_{q^{\star}, t+\Delta} = 0$ and the contribution
to the miner's Brier component is $(0.07 - 0)^2 = 0.0049$. A miner that had instead
emitted $s_i = 0.9$ on the same query would contribute $(0.9 - 0)^2 = 0.81$ --- two
orders of magnitude worse --- and would be ranked correspondingly lower by every
validator that observed the same query, which would clip its consensus weight
$\overline{W}_j$ via Eq.~\eqref{eq:yuma-consensus}. The same machinery applies in the
opposite direction: on a query whose realised outcome is $E = 1$, the miner that emitted
$s_i = 0.07$ contributes $0.86$ to its Brier and is clipped down. This is the operational
form of hypotheses H1 and H2 in Section~\ref{sec:rq}.

\section{Miner Protocol Specification}\label{sec:miner}

A miner $i$ is a thin wrapper over the OmniRisk API, exposed at the Bittensor subnet's
miner endpoint. For each route query $q$ as defined in Eq.~\eqref{eq:route}, the miner
returns a tuple $(s_i, c_i, r_i) \in [0,1] \times [0,1] \times \mathcal{R}$, where
$s_i$ is its estimate of the route risk score $s(q,t)$ from Eq.~\eqref{eq:risk}
($s_i = 0$ denoting nominal and $s_i = 1$ denoting an unrouteable lane), $c_i$ is the
miner's self-reported confidence (Section~\ref{sec:formal:calibration}), and $r_i$ is a
structured reasoning record listing which on-chain observations were used, which DEX
quotes were consulted, and which historical incidents from the BridgeGuard
corpus~\cite{wu2025safeguarding} were the closest analogues.

\paragraph{Determinism constraint.} Yuma scoring requires that two miners using the
same internal logic on the same query at the same block height return the same answer.
Otherwise validators cannot meaningfully aggregate them, and good miners are penalised
for irreducible randomness rather than for being wrong. We pin this by (i)~quantising
$s$ and $c$ to a fixed precision (we use 1/256), (ii)~anchoring the miner to a specific
chain head (a 32-byte block hash per chain consulted), and (iii)~specifying the
ordering and version of any external data sources in $r$. This is the same discipline
that makes optimistic-ML rollup systems
~\cite{conway2024opml} workable: outputs must be a deterministic function of the
declared inputs.

\section{Validator Mechanism}\label{sec:validator}

A validator queries a subset of miners on a stratified sample of routes, scores each
response against the realised-event ground truth defined in
Section~\ref{sec:formal:realised}, and submits a Yuma weight vector
$\mathbf{w} \in [0,1]^{N}$ over the $N$ miners it observed. The on-chain Yuma rule then
aggregates $\mathbf{w}$ across all subnet validators using a stake-weighted threshold and
emits rewards~\cite{opentensor2024yuma}.

The scoring rule is the validator loss $L_i$ given in Eq.~\eqref{eq:vloss}. It has three
components, each grounded in a precise definition from Section~\ref{sec:formal}:
\begin{enumerate}
  \item \textbf{Realised-event accuracy} (the $\alpha$ term, Eq.~\eqref{eq:brier}).
        For routes scored between $t$ and $t + \Delta$, the miner's score $s_i$ is
        evaluated against the realised-loss indicator $E_{q, t+\Delta}$
        (Section~\ref{sec:formal:realised}), which is the disjunction of failed
        execution, abnormal slippage, bridge/lane delay, and exploit-cluster association
        from the BridgeGuard corpus~\cite{wu2025safeguarding}. We deliberately exclude
        downstream channels such as ``post-route drawdown'' that would conflate
        route-execution quality with the user's holding decision.
  \item \textbf{Consistency} (the $\beta$ term). If the same query is issued at $t$ and
        $t + \delta$ for small $\delta$ during which the chain state is materially
        unchanged in the precise sense of Section~\ref{sec:formal:material}, the
        responses must be byte-identical (modulo the quantisation precision of
        Section~\ref{sec:miner}). Inconsistency under no material chain-state change
        indicates either non-determinism (already penalised per Section~\ref{sec:miner})
        or unstable internal state.
  \item \textbf{Confidence calibration} (the $\gamma$ term). Calibration is measured by
        the expected calibration error over reliability bins of $c_i$, with the precise
        definition in Section~\ref{sec:formal:calibration}. Miners that report $c_i = 1$
        on every query are automatically miscalibrated, and miners whose self-reported
        confidence tracks realised accuracy receive lower $L_i$ and therefore higher
        Yuma weight, which is the operational form of hypothesis H2
        (Section~\ref{sec:rq}).
\end{enumerate}

The key design property is that validators score the \emph{quality} of intelligence, not
just its presence. Miners that route every score through the same underlying OmniRisk
API but apply different downstream models can still be differentiated, because the
realised-event channel breaks ties.

\section{Cross-Chain Risk Model}\label{sec:risk}

We formalise three risk axes that emerge \emph{between} chains rather than within them:
\begin{description}
  \item[Liquidity fragmentation.] The same token has independent depth on each chain;
        the worst-of distribution dominates a multi-hop route. The empirical
        misallocation documented by~\cite{gogol2024liquidity} for WETH--USDC pools is
        the simplest instance of this; the OFT lane fabric of LayerZero~V2 generalises
        the problem to arbitrary fungible tokens~\cite{layerzero2024oft}.
  \item[Bridge / route dependency risk.] A given OFT lane inherits the security
        configuration of its DVN set and Executor~\cite{zarick2024layerzero}. The
        bridge-security literature catalogues how individual configuration choices map
        onto exploited vulnerability classes
        ~\cite{zhang2023sok,augusto2024sok,wu2025safeguarding}.
  \item[Anomaly propagation.] An exploit on one chain can spread to others via shared
        bridges, shared liquidity, and shared dependencies. Multichain fraud-detection
        datasets~\cite{palaiokrassas2023multichain} provide the empirical basis for
        modelling this, but the cross-chain dimension remains under-instrumented in
        published work.
\end{description}
The unifying observation is that cross-chain risk is \emph{a property of routes, not of
tokens or chains in isolation}, and a verification layer that scores claims about routes
is therefore the natural primitive.

\section{Subtensor EVM (Chain~ID~945) Integration}\label{sec:subtensor}

The Subtensor EVM is an Ethereum-compatible execution environment that runs inside the
Subtensor Substrate runtime, with TAO as the gas asset~\cite{opentensor2024evm}. It is
not just another EVM lane for RayaChain; it is the execution environment whose
validator set is already economically aligned to scoring the quality of off-chain
intelligence, by virtue of the Yuma mechanism~\cite{opentensor2024yuma,
steeves2022incentivizing}. We use it for two distinct purposes:
\begin{enumerate}
  \item \textbf{Anchoring of risk-score commitments.} Each batch of miner outputs is
        committed as a Merkle root on Subtensor EVM, so a downstream contract on any
        of the seven existing RayaChain lanes can fetch a proof that a given
        $(s, c, r)$ tuple was produced and made available at a specific block height,
        prior to the route being taken.
  \item \textbf{Anchoring of validator-consensus outputs.} The aggregated Yuma reward
        vector for the cross-chain risk subnet is itself published on Subtensor EVM,
        which gives a downstream auditor a single canonical record of which miners
        were trusted at any given epoch.
\end{enumerate}
This aligns the economic incentives of the AI layer (miners and validators), the
liquidity layer (RayaChain LPs), and the verification layer (TAO holders staking on
the subnet).

\section{Data and Evaluation Framework}\label{sec:eval}

The evaluation we are building rests on three datasets:
\begin{enumerate}
  \item \textbf{RayaChain flow logs.} OFT send/receive events across the seven live
        lanes, with per-lane metadata (DVN set, Executor, observed latencies,
        message-failure counts).
  \item \textbf{Realised-event labels.} A union of (a)~slippage and execution-failure
        outcomes derived from the flow logs (the $E1$--$E3$ channels of
        Section~\ref{sec:formal:realised}) and (b)~external incident labels from
        peer-reviewed and grey-literature corpora (the $E4$ channel); the BridgeGuard
        dataset~\cite{wu2025safeguarding} (203 attack transactions across
        49~incidents) is one of the curated bases.
  \item \textbf{Replay routes.} For each historical route observed in (1), we replay it
        to the OmniRisk-backed miners under the configuration at the corresponding block
        height (Section~\ref{sec:miner}), and score against (2).
\end{enumerate}
The headline metrics are realised-event ROC-AUC, precision-recall AUC (the latter is
required because the BridgeGuard prior gives a class-imbalance ratio of approximately
$1{:}197$ for $E4$ events and ROC-AUC alone is misleading at that imbalance,
cf.~Section~\ref{sec:related:bridge-detection}), early-detection lead time (how many
blocks before an event the score crossed threshold), expected calibration error
(Section~\ref{sec:formal:calibration}), and Yuma reward retention (a miner's reward share
over a window normalised by its scored share, as a measure of alignment with the
validator consensus, computed against the consensus weight $\overline{W}_j$ of
Eq.~\eqref{eq:yuma-consensus}).

\subsection{Baselines}\label{sec:eval:baselines}

A claim that a verifiable cross-chain risk primitive outperforms the status quo is only
defensible against explicit baselines. Each of the following is implemented as a
\emph{reference miner} on the same subnet, scored under the same validator loss
$L_i$ as any production miner, and reported alongside the OmniRisk miner in every
follow-up evaluation. They span the design space from naive to credible analytic
alternatives.

\begin{description}
  \item[B1: Random / naive baseline.] Returns $s = 0.5$ on every query, with
        $c = 0.5$. This is the lower bound: any miner whose Brier score does not improve
        on B1 is providing no information.
  \item[B2: Single-chain liquidity-only model.] A logistic regressor over destination-chain
        liquidity depth, recent destination-chain volatility, and asset identifier ---
        \emph{no source-chain or lane features}. This baseline isolates the value of
        cross-chain telemetry; rejection of H1 (Section~\ref{sec:rq}) corresponds to the
        OmniRisk miner failing to improve early-detection lead time over B2.
  \item[B3: Static bridge-risk heuristic.] A fixed mapping from
        $(c_s, c_d, \text{DVN set}, \text{Executor})$ to a constant risk tier, calibrated
        once at evaluation start and never updated. This baseline isolates the value of
        the agentic, adaptive component of OmniRisk against a static analyst-tier
        product.
  \item[B4: Generic anomaly detector.] A standard isolation-forest or one-class SVM
        trained on the union of source- and destination-chain features but without any
        bridge-/lane-specific structure or the BridgeGuard incident corpus. This baseline
        isolates the value of incident-grounded supervision against unsupervised
        anomaly scoring.
\end{description}

For each baseline we report (i)~Brier score, (ii)~realised-event AUC, (iii)~early-detection
lead time, (iv)~expected calibration error, and (v)~Yuma reward retention over the
replay window. The OmniRisk miner is required to dominate B1 strictly on all five
metrics, dominate B2 on (iii) for hypothesis H1 to survive, and dominate B3 on (i)--(iii)
for the agentic-vs-static claim of Section~\ref{sec:compare} to be defensible.

\section{Economic Model}\label{sec:economics}

The economic flows are:
\begin{itemize}
  \item Miners are rewarded in TAO via Yuma emissions for risk outputs that the
        validator consensus accepts~\cite{opentensor2024yuma,steeves2022incentivizing}.
  \item Validators are rewarded for assigning weights that align with the consensus,
        which is the natural Yuma incentive to score accurately.
  \item OmniRisk is the infrastructure provider: it operates the underlying analytics
        platform that miners wrap, and it monetises this through its existing API
        surface to non-Bittensor consumers.
\end{itemize}
We avoid claims about token-economic equilibria here. The Bittensor critical-analysis
literature~\cite{lui2025bittensor} documents real concerns about Bittensor's
centralisation properties under dynamic TAO, and any quantitative economic claim about
this subnet should be revisited against that work in a follow-up.

\section{Security Model}\label{sec:security}

The threat surface is dominated by three classes:
\begin{description}
  \item[Adversarial miners.] A miner can attempt to bias the risk score it emits, for
        example to suppress a warning on a lane it benefits from. The Yuma scoring rule
        bounds the influence of such a miner to its share of stake-weighted
        rank~\cite{steeves2022incentivizing}; the determinism and consistency
        components of the validator rule (Section~\ref{sec:validator}) further limit
        steady-state bias.
  \item[Manipulated liquidity signals.] An attacker can briefly distort the on-chain
        signals that miners ingest --- e.g.\ wash-trading a thin pool to mask an exit.
        Our mitigation is multi-source data validation: the validator's
        realised-event channel (Section~\ref{sec:validator}) is by construction
        retrospective and ignores manipulated quote-time signals.
  \item[Cross-chain spoofing.] A malicious bridge can claim a state on one chain that
        is not consistent with the source. The OFT design~\cite{layerzero2024oft}
        and the V2 DVN model~\cite{zarick2024layerzero} reduce, but do not eliminate,
        this risk; the bridge-security literature documents the residual surface in
        depth~\cite{zhang2023sok,augusto2024sok,wu2025safeguarding}.
\end{description}

\section{Implementation Status}\label{sec:status}

We separate what is deployed from what is in design (Table~\ref{tab:status}).

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Component} & \textbf{Status as of 2026-05} & \textbf{Reference} \\
\midrule
OmniRisk intelligence layer        & Live in production       & this paper \\
RayaChain lanes (7 chains)         & Live / actively deploying & \cite{zarick2024layerzero,layerzero2024oft} \\
Subtensor EVM lane (chain~945)     & Planned (8th lane)        & \cite{opentensor2024evm} \\
Cross-chain risk subnet (Bittensor)& Design proposal           & this paper \\
Validator scoring rule             & Design proposal           & \S\ref{sec:validator} \\
Score-commitment anchoring on TAO  & Design proposal           & \S\ref{sec:subtensor} \\
\bottomrule
\end{tabular}
\caption{Implementation status. Components marked \emph{design proposal} are
specified in this paper but not yet deployed, and quantitative claims about them are
deferred to follow-up work.}
\label{tab:status}
\end{table}

\section{Roadmap}\label{sec:roadmap}

The work plan implied by Table~\ref{tab:status} is, in order:
(i)~finalise the miner API and the determinism contract (Section~\ref{sec:miner});
(ii)~stand up the cross-chain risk subnet on Subtensor and integrate the validator
scoring rule (Section~\ref{sec:validator});
(iii)~deploy the score-commitment contracts on Subtensor EVM (Section~\ref{sec:subtensor});
(iv)~ship the eighth RayaChain lane into Subtensor EVM;
(v)~publish the evaluation results from Section~\ref{sec:eval} as a follow-up paper.

\section{Comparative Advantage}\label{sec:compare}

Existing cross-chain risk tooling is some combination of single-chain, static, or
black-box. The architecture we propose is differentiated on three axes simultaneously:
\begin{itemize}
  \item \textbf{Agentic, not static.} The intelligence layer adapts its tool use to the
        question, rather than precomputing a fixed risk surface.
  \item \textbf{Cross-chain-native, not single-chain.} Routes, not tokens, are the
        primitive being scored, which is the level at which fragmentation and
        bridge-dependency risk actually emerge~\cite{gogol2024liquidity,
        zhang2023sok}.
  \item \textbf{Verifiable, not black-box.} The Bittensor verification layer makes
        miner outputs economically expensive to bias, and the on-chain commitments on
        Subtensor EVM (Section~\ref{sec:subtensor}) make their history auditable from
        any of the seven other lanes.
\end{itemize}

\section{Limitations}\label{sec:limitations}

We explicitly flag the following limitations.
\begin{itemize}
  \item \textbf{Most quantitative claims are deferred.} The score-commitment
        contracts, the Bittensor subnet, and the evaluation framework
        (Section~\ref{sec:eval}) are design proposals; this paper does not yet report
        AUC numbers, reward-retention curves, or measured comparisons against the
        baselines of Section~\ref{sec:eval:baselines}. The hypotheses H1 and H2
        (Section~\ref{sec:rq}) are stated as falsifiable claims, not as established
        results.
  \item \textbf{Bittensor centralisation risk.} Independent empirical work has raised
        concerns about Bittensor's actual decentralisation under the dynamic-TAO
        regime~\cite{lui2025bittensor}. We treat the Yuma collusion-resistance bound
        of \cite{steeves2022incentivizing} as a \emph{design property of the algorithm}
        rather than as a guarantee about the deployed network's stake distribution.
  \item \textbf{Withdrawn primary source.} The original BitTensor preprint
        (arXiv:2003.03917) was withdrawn by its authors as obsolete on 2021-11-10. Any
        prior whitepaper that cited it should be reread against the canonical
        \cite{rao2021bittensor} or the workshop
        formalisation~\cite{steeves2022incentivizing}.
  \item \textbf{Determinism in agentic LLM components.} Section~\ref{sec:miner}'s
        determinism contract is straightforward for classical analytics but is more
        delicate when the miner uses an LLM under the hood; the
        consistency component of the validator rule
        (Section~\ref{sec:validator}) is what catches the residual non-determinism in
        practice.
\end{itemize}

\section{Conclusion}\label{sec:conclusion}

Cross-chain risk has settled into a regime where the dominant losses come from a small
number of well-understood failure modes in the bridge layer, but where the risk
\emph{intelligence} that participants need is itself untrusted. We have presented an
architecture that splits intelligence (OmniRisk), data and execution (RayaChain on
LayerZero~V2), and verification (Bittensor's Yuma Consensus, anchored on Subtensor EVM
chain~ID~945) into three layers with clean economic boundaries. The result, when fully
deployed, is the first production cross-chain risk primitive whose outputs are
verifiable in the same sense that the underlying transfers are: by an independent,
economically aligned, decentralised process. The intermediate deliverables on the
roadmap (Section~\ref{sec:roadmap}) are testable in isolation, and we will report on
each in follow-up work.

\section*{Acknowledgements}
We thank the OmniRisk engineering team for the production telemetry that grounds
Section~\ref{sec:eval}, and the Opentensor Foundation maintainers for the Subtensor
EVM testnet documentation that made the Section~\ref{sec:subtensor} integration plan
concrete. Any errors are the authors'.

\section*{Reproducibility Note}
The bibliography for this paper is the master \texttt{references.bib} of the
\texttt{omnirisk-papers} repository. Every entry was reverified against its primary
source on 2026-05-03 prior to inclusion. Two entries were added in revision v0.4 to
fill the related-work gaps flagged in academic review (\cite{ganescu2024trust} for
zero-knowledge ML verification, \cite{werner2022sok} for the broader DeFi oracle
landscape) and were verified against their arXiv abstract pages on the same date.
Revision v0.5 deepened engagement with three previously cited primary sources by
extracting concrete material rather than relying on summary attribution: the Yuma
Consensus equations of Eqs.~\eqref{eq:yuma-consensus} and~\eqref{eq:vweight} were
extracted from~\cite{opentensor2024yuma}; the \texttt{UlnConfig} and
\texttt{ExecutorConfig} field names of Section~\ref{sec:formal:material} from the
LayerZero V2 developer documentation~\cite{layerzero2024oft,zarick2024layerzero}; the
$1{:}197$ class-imbalance ratio of Section~\ref{sec:related:bridge-detection} from the
BridgeGuard dataset described in~\cite{wu2025safeguarding}. The retracted preprint
arXiv:2003.03917 is intentionally excluded; see the comment block in
\texttt{references.bib} and Section~\ref{sec:limitations} for the rationale.

\printbibliography

\end{document}
