\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabularx}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage[backend=biber,style=numeric-comp,sorting=none]{biblatex}

\setlength{\emergencystretch}{3em}
\widowpenalty=10000
\clubpenalty=10000
\AtBeginBibliography{\sloppy\raggedright\setlength{\emergencystretch}{6em}}

\addbibresource{references.bib}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=blue!50!black,
  urlcolor=blue!50!black,
  pdftitle={A Cost-Aware, Self-Adaptive Agentic AI Engine for Risk Intelligence in the Internet of Value},
  pdfauthor={Basel Magableh; OmniRisk Research},
}

\title{A Cost-Aware, Self-Adaptive Agentic AI Engine\\
       for Risk Intelligence in the Internet of Value\\[0.2em]
       \large An engineering case study}
\author{%
  Basel Magableh\textsuperscript{1,*} \quad\quad OmniRisk Research\textsuperscript{2} \\[0.8em]
  \normalsize\textsuperscript{1}School of Computer Science, Technological University Dublin, Ireland \\
  \normalsize\textsuperscript{2}Rayachain Lab, Dublin, Ireland \\[0.4em]
  \normalsize\textsuperscript{*}Correspondence: \texttt{basel.magableh@tudublin.ie}
}
\date{May 2026 \quad (draft v0.2 --- targeting Wiley \emph{Software: Practice and Experience})}

\begin{document}
\maketitle

\begin{abstract}
Agentic AI systems are now routinely deployed in heterogeneous,
partially-trusted environments where the marginal risk that matters is
\emph{composite}: route, sentiment, liquidity, and the policy a system
is willing to commit to. Existing software-engineering experience for
machine-learning systems~\cite{sculley2015hidden,amershi2019software}
addresses an offline-train-then-deploy regime, and recent surveys of
LLM-based autonomous agents~\cite{wang2024survey} catalogue what such
agents are made of, but neither tells a practitioner how to engineer one
that \emph{stays cheap and stays calibrated} for months under adversarial
inputs. This paper reports the engineering --- not the algorithms --- of
the OmniRisk system, an agentic risk-intelligence platform deployed on
the blockchain Internet of Value. We describe two engineering
contributions.
\textbf{(C1) A cost-aware AI engine}: a prediction router and inference
cascade in the lineage of FrugalGPT~\cite{chen2023frugalgpt} and
RouteLLM~\cite{ong2024routellm}, extended with a decentralised-economic
verifier so that ``cost'' includes not only inference dollars but the
expected social cost of trusting a single model.
\textbf{(C2) A self-adaptive agentic AI}: a MAPE-K
loop~\cite{salehie2009self,weyns2020introduction} in which the LLM-
mediated agent observes its own miss-rate, re-weights its prediction
ensemble, and re-tightens its constitutional action
constraints~\cite{bai2022constitutional} on a versioned calibration
manifest, with a non-negotiable honesty discipline (publish denominators,
source-of-truth artefacts, no silent retraining).
We report two operational case studies as evidence that the two
contributions work in production: a 27-hour policy-constrained Solana
liquidity stress experiment and a 168-hour prediction-router calibration
arc, the latter reported with explicit class-imbalance caveats. The
contribution is methodological: a description of the discipline by which
a cost-aware, self-adaptive agentic AI system was assembled, calibrated,
monitored, and made falsifiable.
\end{abstract}

\medskip
\noindent\textbf{Keywords ---} software engineering for AI; cost-aware
inference; self-adaptive software; MAPE-K; agentic systems; LLM
operations; risk intelligence; case study.

\section{Introduction}\label{sec:intro}

The software-engineering literature on machine-learning-intensive
systems has, over a decade, settled an uncomfortable consensus: most of
the cost is not in the model. Sculley
et~al.~\cite{sculley2015hidden} catalogued the system-level burdens
(entanglement, hidden feedback loops, undeclared consumers, configuration
debt, monitoring debt) that ML systems pay over and above ordinary code.
Amershi et~al.~\cite{amershi2019software} documented at industrial scale
the workflow discontinuities that ML imposes on traditional SE practice.
The recent emergence of \emph{agentic} systems --- LLM-mediated software
components that select and execute actions, not merely produce
predictions~\cite{wang2024survey,bai2022constitutional} --- intensifies
each of those burdens, because the agent is now also a closed-loop
controller whose decisions feed back into the data it later consumes.
Two concerns then dominate practice: \emph{cost} (every action and every
inference is metered) and \emph{adaptation} (the environment shifts
faster than any offline-trained model can keep up with).

This paper concerns the engineering of one such system, deployed in a
genuinely adversarial setting. The OmniRisk system monitors and reasons
about risk on the blockchain ``Internet of Value'' (IoV): a heterogeneous
network of chains, bridges, liquidity venues, and off-chain narratives in
which the dominant marginal risk to any participant is
\emph{composite}~\cite{augusto2024sok,zhang2023sok,belchior2021survey,wu2025safeguarding},
not a property of any single chain. We deliberately set aside the
security and finance arguments (treated in a companion
paper~\cite{omnirisk2026defending}) and ask the SE practitioner's
question instead: \emph{how do you keep an agentic AI system that
forecasts and acts on a partially-trusted environment both cheap and
calibrated, for months, in a way that is reproducible and reviewable?}

\paragraph{Contributions.}
The paper makes two engineering contributions and supplies operational
evidence for each.

\begin{description}
  \item[(C1) A cost-aware AI engine]
        (\S\ref{sec:cost-aware}). A prediction router and inference
        cascade in the lineage of FrugalGPT~\cite{chen2023frugalgpt} and
        RouteLLM~\cite{ong2024routellm}, extended in two ways. First, it
        composes inference-dollar cost with an \emph{economic-trust}
        cost obtained by routing high-stakes inferences through a
        decentralised verification subnet so that ``cheap and wrong''
        is priced as expensive. Second, the constitutional
        agent~\cite{bai2022constitutional} sits under an explicit token
        budget at every step, so that adversarial prompts cannot run the
        system out of money.
  \item[(C2) A self-adaptive agentic AI]
        (\S\ref{sec:self-adaptive}). A MAPE-K-style adaptation
        loop~\cite{salehie2009self,weyns2020introduction} in which the
        agent observes its own miss-rate against a frozen calibration
        manifest~\cite{omnirisk2026calibration,omnirisk2026predmetrics},
        re-weights the prediction ensemble when drift is detected, and
        re-tightens its constitutional action constraints when the
        verifier disagrees. The loop is made trustworthy by an
        engineering honesty discipline (publish denominators,
        source-of-truth artefacts, no silent retraining) which we treat
        as the most transferable contribution of this paper.
\end{description}

\paragraph{Operational evidence.}
We support both contributions with two case studies on real production
artefacts: a 27-hour Solana liquidity stress
experiment~\cite{omnirisk2026soak,omnirisk2026deploy} and a 168-hour
prediction-router calibration
arc~\cite{omnirisk2026predmetrics,omnirisk2026calibration} reported with
explicit class-imbalance caveats. The case studies are operational
rather than controlled, and we report their limits honestly.

\paragraph{Companion paper and scope.} A companion
manuscript~\cite{omnirisk2026defending} treats the same system from a
security-and-verification angle (threat model, validator-loss
decomposition, formal guarantees). The present paper is the
\emph{software-engineering} cut: composition, cost, adaptation,
operability, and the discipline of being calibrated. Where security
facts are required, we cite the companion paper rather than restate
them.

\section{Background and related work}\label{sec:background}

\subsection{SE for ML and agentic systems}\label{sec:bg:se-ml}

Sculley et~al.~\cite{sculley2015hidden} and
Amershi et~al.~\cite{amershi2019software} together set the SE-for-ML
agenda: ML systems pay system-level costs that ordinary SE practice does
not anticipate, and the productive engineering response is to make the
ML workflow first-class. Subsequent work has extended this to LLM-
mediated agentic systems~\cite{wang2024survey}, which add two
qualitatively new concerns: (i) the agent's outputs are \emph{actions}
that perturb the environment from which its next inputs are drawn, and
(ii) the agent is constituted in part by natural-language constraints
(``constitutions'') rather than only by code~\cite{bai2022constitutional}.
The agent decomposition we adopt --- profile, memory, planning, action ---
follows Wang et~al.'s unified
framework~\cite{wang2024survey} and is the vocabulary of
\S\ref{sec:self-adaptive}.

\subsection{Cost-aware LLM serving}\label{sec:bg:cost}

Two recent threads in the LLM-serving literature directly motivate our
cost-aware design. FrugalGPT~\cite{chen2023frugalgpt} introduces an LLM
\emph{cascade} that routes a query through models of increasing cost
until a confidence threshold is met, reducing per-query inference cost
without (in their reported settings) reducing quality. RouteLLM~\cite{ong2024routellm}
trains a learned router over a pool of models on human-preference data so
that easy queries go to small/cheap models and hard ones are escalated.
We treat both as the canonical pattern for cost-aware inference and
generalise it in \S\ref{sec:cost-aware} by composing it with an
\emph{economic-trust} cost paid through a decentralised verification
substrate~\cite{rao2021bittensor,steeves2022incentivizing,opentensor2024yuma,conway2024opml,ganescu2024trust}.

\subsection{Self-adaptive software}\label{sec:bg:self-adaptive}

Self-adaptive systems are the older intellectual lineage that
anticipated much of what agentic AI now demands. Salehie and
Tahvildari's survey~\cite{salehie2009self} formalises the MAPE-K
(Monitor--Analyse--Plan--Execute over Knowledge) feedback loop as the
canonical decomposition of a self-adaptive system, and Weyns'
textbook~\cite{weyns2020introduction} extends MAPE-K with explicit
treatment of uncertainty management, feedback control, and online
learning. The self-adaptive AI of \S\ref{sec:self-adaptive} is a
particular instantiation of MAPE-K in which the Analyse and Plan stages
are LLM-mediated and the Execute stage is policy-constrained.

\subsection{Risk intelligence in the Internet of Value}\label{sec:bg:iov}

We summarise the IoV setting briefly; it motivates the case studies
without being the object of this paper. Cross-chain bridges are now both
the dominant capital route and the dominant exploit
surface~\cite{augusto2024sok,wu2025safeguarding,chainalysis2022bridges};
liquidity is fragmented across chains and venues so that the same
logical asset can be priced and routed
inconsistently~\cite{zhang2023sok,palaiokrassas2023multichain,gogol2024liquidity};
sentiment shocks propagate at a faster cadence than chain-state changes,
so a monitor that ignores text drift is blind to events whose on-chain
footprint is downstream of the
narrative~\cite{schumaker2009textual,xing2018natural,day2016deep,zhang2025sentistack};
and an LLM-mediated agent that selects and executes on-chain actions
introduces a closed-loop risk that single-chain monitors do not
capture~\cite{wang2024survey,bai2022constitutional}.

\section{(C1) The cost-aware AI engine}\label{sec:cost-aware}

The system's first engineering contribution is to treat \emph{cost} as a
first-class quantity composed of three terms: \emph{inference cost}
(dollars and tokens spent on model calls), \emph{economic-trust cost}
(the price of trusting a single model on a high-stakes decision), and
\emph{action cost} (the on-chain or operational price of executing a
decision). The engine that manages these three together is composed of
three cooperating gates.

\subsection{Gate~A: a confidence-cascading prediction router}\label{sec:cost-aware:cascade}

Following the cascade pattern of FrugalGPT~\cite{chen2023frugalgpt} and
the learned-router pattern of RouteLLM~\cite{ong2024routellm}, the
prediction router holds a small ensemble of forecasters of varying cost
and quality (cheap heuristic baselines, mid-tier ML models, and
expensive ensembles or LLM-mediated calls). Each query is first answered
by the cheapest forecaster; the answer is accepted if its
confidence-interval width and its agreement with a frozen calibration
manifest~\cite{omnirisk2026calibration} are both within a per-symbol
threshold. Otherwise it is escalated. The engineering point is not the
specific cascade --- that is a known
pattern~\cite{chen2023frugalgpt,ong2024routellm} --- but that the
acceptance threshold is itself a versioned, observable artefact, and
that the router publishes its own escalation rate and miss rate alongside
its forecasts so that downstream consumers can audit the cascade's
discipline.

\subsection{Gate~B: economic-trust cost via decentralised
verification}\label{sec:cost-aware:verification}

For high-stakes inferences --- those that the scenario engine of
\S\ref{sec:cost-aware:scenario} would permit to drive an on-chain
action --- the router's accepted output is routed through a Bittensor-
style verification
subnet~\cite{rao2021bittensor,steeves2022incentivizing,opentensor2024yuma,opentensor2024evm,lui2025bittensor},
in which independently incentivised validators score the inference. The
engineering insight is to read this as a \emph{cost} term: the system
pays validators to disagree with the router when the router is wrong, so
that ``cheap and wrong'' is priced as expensive. This is methodologically
adjacent to opML~\cite{conway2024opml,ganescu2024trust} but treated here
as one term in the cost ledger rather than as a security primitive (for
which see~\cite{omnirisk2026defending}). When the verification subnet
disagrees with the router's output, the system fails closed and refuses
the action, which is the desired behaviour from a cost-aware standpoint:
the most expensive decision is the one taken on a wrong inference.

\subsection{Gate~C: action-cost gating in the agentic engine}\label{sec:cost-aware:scenario}

The constitutional agent~\cite{bai2022constitutional} that maps
prediction-and-sentiment state to candidate actions sits under an
explicit token budget at every step, and emits each candidate as a
pre-committed scenario-conditional API call (an \emph{action program})
in the sense of Glasserman's scenario-based Monte
Carlo~\cite{glasserman2003monte}. The action program is a reviewable
artefact: a human or another system can audit it before any on-chain
execution. From a cost-aware perspective, the discipline is that
action-cost is paid only when (Gate~A) the cascade has accepted a
forecast at the appropriate cost tier and (Gate~B) the verifier has not
disagreed with that forecast.

\subsection{The cost ledger}\label{sec:cost-aware:ledger}

The three gates together produce, per decision, an explicit cost ledger
of the form
\(\mathrm{cost}(d) = c_{\text{inf}}(d) + c_{\text{trust}}(d)
+ c_{\text{act}}(d),\)
each term observable, each term tied to a versioned artefact (the
calibration manifest~\cite{omnirisk2026calibration,omnirisk2026predmetrics},
the validator scores, and the action program). We do not claim a new
algorithm; we claim that engineering each term as an independently
versioned and observable quantity is the discipline that makes the
overall system cost-auditable.

\section{(C2) Self-adaptive agentic AI}\label{sec:self-adaptive}

The system's second engineering contribution is a self-adaptive
loop~\cite{salehie2009self,weyns2020introduction} that closes around the
cost-aware engine of \S\ref{sec:cost-aware}, so that the agent
\emph{observes its own miss-rate and reconfigures itself} on a versioned
calibration manifest rather than waiting for an offline retraining cycle.

\subsection{A MAPE-K instantiation}\label{sec:self-adaptive:mapek}

Self-adaptation is structured as MAPE-K~\cite{salehie2009self}:
\begin{description}
  \item[Monitor.]
        Sentiment fusion~\cite{macrocosmos2024datauniverse,birdeye2026docs,dexscreener2026docs,yang2023fingpt,wu2023bloomberggpt,gandhi2023multimodal,lai2023multimodal,long2019deep,zhang2025sentistack}
        and the read-side of the prediction
        router~\cite{omnirisk2026predmetrics} together produce a stream
        of (forecast, realised-outcome, validator-score) triples.
  \item[Analyse.]
        The verification subnet (\S\ref{sec:cost-aware:verification})
        and the calibration
        manifest~\cite{omnirisk2026calibration} together compute a
        running residual: how badly is the cascade missing, on which
        symbols, with what class imbalance.
  \item[Plan.]
        The agentic engine~\cite{wang2024survey,bai2022constitutional}
        chooses between three remedies --- re-weight the cascade, narrow
        the scenario engine's permitted action set, or escalate to a
        human operator --- using the agent decomposition of
        \emph{profile, memory, planning, action} prescribed in the
        LLM-agent survey~\cite{wang2024survey}.
  \item[Execute.]
        The scenario engine (\S\ref{sec:cost-aware:scenario}) emits the
        adaptation as a versioned change to the cascade thresholds, the
        ensemble weights, or the
        constitution~\cite{bai2022constitutional}, each of which is a
        diffable artefact in
        git~\cite{omnirisk2026systemdesign,omnirisk2026agentic}.
  \item[Knowledge.]
        The shared, versioned state is the calibration
        manifest~\cite{omnirisk2026calibration,omnirisk2026predmetrics},
        the contract-and-route
        registry~\cite{rayachain2026registry}, and the
        constitution~\cite{bai2022constitutional}. The point worth
        emphasising is that Knowledge is \emph{checked in}, not
        latent in a model's weights.
\end{description}
This mapping is non-trivial: each MAPE-K box has a named owning
component, a separately observable failure mode, and a separate test
harness, in the modular spirit that
SE-for-ML~\cite{sculley2015hidden,amershi2019software} prescribes and
that Weyns' textbook~\cite{weyns2020introduction} formalises.

\subsection{An honesty discipline for adaptation}\label{sec:self-adaptive:honesty}

A self-adaptive AI in an adversarial environment is only useful if its
self-reports are trustworthy. We adopt three engineering rules,
non-negotiable in the OmniRisk codebase, and we view them as the most
transferable contribution of this paper.

\paragraph{Rule 1: publish denominators.} Every accuracy or skill metric
the adaptation loop reads is published with the class-imbalance
denominator that produced it. The 168-hour prediction-router calibration
arc reported in~\S\ref{sec:case-studies:router} \emph{drops} when
reported honestly: a 99.34\% \texttt{didCrashAccuracy} on a class-imbalanced
cohort is far less impressive than the headline number suggests, and the
engineering response is to say so on the same page as the
number~\cite{omnirisk2026predmetrics}.

\paragraph{Rule 2: source-of-truth artefacts, not narratives.}
Calibration manifests~\cite{omnirisk2026calibration}, soak
traces~\cite{omnirisk2026soak}, and the
contract-and-route registry~\cite{rayachain2026registry} are committed,
versioned, and citeable. A claim in the manuscript is admissible only if
it points to such an artefact; this is the engineering analogue of
``every release ships with its evidence pack''.

\paragraph{Rule 3: refuse silent retraining.} Models, prompts, ensemble
weights, and constitutions are versioned together. A change to any one
of them is a release event with an audit trail; ``the model improved
overnight'' is a defect, not a feature, because it makes the system's
behaviour irreproducible. Without this rule, MAPE-K \emph{Execute}
becomes a back door that lets the system silently optimise itself away
from the behaviour its operators reviewed.

\section{Operational evidence}\label{sec:case-studies}

We supply two operational case studies that exercise the cost-aware
engine of \S\ref{sec:cost-aware} and the self-adaptive loop of
\S\ref{sec:self-adaptive} on real production artefacts. Neither is a
controlled experiment; both are operational reports on dates and runs
that are now fixed in git, with their evidence packs cited.

\subsection{27-hour Solana liquidity stress experiment}\label{sec:case-studies:soak}

We ran a 27-hour stress experiment on Solana in which the constitutional
agent operated against a single liquidity pool, with the scenario engine
gating every on-chain
action~\cite{omnirisk2026soak,omnirisk2026deploy}. As evidence for C1
(cost-aware engine), the trace shows: every action the agent proposed
was emitted as a reviewable artefact before execution; the verification
subnet (Gate~B) acted as an independent gate that prevented two proposed
actions from executing because it disagreed with the router; and no
inference was permitted to escalate past Gate~A's per-symbol confidence
threshold without being repriced through Gate~B. As evidence for C2
(self-adaptive loop), the agent observed two regime shifts in the
27-hour window and re-weighted the prediction-router cascade in response
to each, with the re-weighting committed as a diff to the cascade
thresholds rather than as a silent ensemble retrain.

\subsection{168-hour prediction-router calibration arc}\label{sec:case-studies:router}

Over a 168-hour window we evaluated the prediction router against the
production metrics
table~\cite{omnirisk2026predmetrics,omnirisk2026calibration}. Reported
honestly, the headline live-API accuracy on the April~2026 Blue API was
approximately 53.15\% --- ``barely above chance'' in the source's own
framing. After the 2026-04-21 horizon-aware crash-probability
calibration fix, the freshest 168-hour cohort (run 2026-05-07
01:41:09~UTC, 915 samples) reports a \texttt{didCrashAccuracy} of
99.34\% and a crash-probability Brier score of 0.1335. The arc must be
read in the context of the class-imbalance and cohort-size caveats
recorded in the source-of-truth manifest: the high accuracy reflects a
heavily skewed prior on the no-crash class, and the engineering value of
reporting it is precisely to expose the gap between the headline metric
and the calibration discipline that produced it. As evidence for C2, the
arc itself is the adaptation: a versioned calibration fix, deployed on a
specific date, traceable in the source-of-truth manifest, and reported
without retroactive narrative-smoothing.

\section{Engineering lessons}\label{sec:lessons}

We claim five lessons that we believe generalise beyond the blockchain
setting to any agentic AI system whose inputs and actions share a
partially-trusted environment.

\begin{enumerate}
  \item \emph{Treat cost as a composition, not a number.} Inference
        cost, economic-trust cost, and action cost are different
        currencies; conflating them is what produces the ``cheap and
        wrong'' failure mode that dominates production agent
        incidents~\cite{chen2023frugalgpt,ong2024routellm}.
  \item \emph{Make the verifier independent.} An economic verifier that
        is co-located with the optimiser conspires with it. Pay an
        independent judge --- the engineering of trust is a
        cost-budget concern, not only a security
        concern~\cite{rao2021bittensor,steeves2022incentivizing}.
  \item \emph{Self-adapt on a versioned manifest, not on weights.}
        MAPE-K \emph{Execute} should write to a diffable artefact; if
        it cannot be reviewed, it cannot be trusted~\cite{salehie2009self,weyns2020introduction}.
  \item \emph{Decompose the agent the way the LLM-agent survey
        prescribes.} Profile, memory, planning, action~\cite{wang2024survey}
        are first-class subcomponents, each independently versioned and
        testable.
  \item \emph{Publish the denominators.} The honesty rules of
        \S\ref{sec:self-adaptive:honesty} are the cheapest engineering
        investment with the highest return on calibration-trust.
\end{enumerate}

\section{Threats to validity and limitations}\label{sec:limits}
% Companion-paper bib notes referenced sibling-manuscript labels that do not
% exist here (sec:case-study, sec:formal:vloss in the IEEE-TIFS cut). Anchor
% them invisibly so the bibliography note macros resolve and the build is
% warning-free.
\phantomsection\label{sec:case-study}%
\phantomsection\label{sec:case-study:prediction-router}%
\phantomsection\label{sec:formal:vloss}%

The case studies of \S\ref{sec:case-studies} are operational rather than
controlled experiments: they report what the system did under real
conditions on dates that are now fixed in
git~\cite{omnirisk2026soak,omnirisk2026predmetrics}. They are not
randomised, do not estimate counterfactuals, and do not generalise
beyond the chains, pools, and time-windows they describe. The
cost-aware composition of \S\ref{sec:cost-aware} reuses the cascade
pattern of FrugalGPT~\cite{chen2023frugalgpt} and the learned-router of
RouteLLM~\cite{ong2024routellm}; we do not claim those primitives as new
contributions, only their composition with an economic-trust cost and an
explicit action-cost gate. The self-adaptive loop of
\S\ref{sec:self-adaptive} is a particular instantiation of MAPE-K in the
sense of Salehie and Tahvildari~\cite{salehie2009self} and
Weyns~\cite{weyns2020introduction}; a comparative study against
alternative decompositions (a single end-to-end LLM agent with no
separate verifier, for instance) is left to future work. The honesty
rules of \S\ref{sec:self-adaptive:honesty} are claims about practice
that we follow in this codebase but cannot prove transfer; replication
on an unrelated system would test that claim.

\section{Conclusion}\label{sec:conclusion}

We have described the engineering --- not the algorithms --- of an
agentic, context-aware AI system that operates in a genuinely
adversarial environment. The contribution is twofold: a \emph{cost-aware
AI engine} that composes inference cost, economic-trust cost, and
action cost into a single observable
ledger~\cite{chen2023frugalgpt,ong2024routellm,rao2021bittensor}; and a
\emph{self-adaptive agentic AI} that closes a MAPE-K loop around that
ledger on a versioned calibration manifest with a non-negotiable
honesty discipline~\cite{salehie2009self,weyns2020introduction,wang2024survey,bai2022constitutional}.
Two operational case studies supply the falsifiable evidence. The
lessons are deliberately general: any agentic AI system whose inputs and
actions share a partially-trusted environment will benefit from making
its verifier independent, its constitution diffable, its denominators
public, and its actions pre-committed.

\printbibliography

\end{document}
